---
title: "Project 2"
author: "Munkhnaran Gankhuyag"
date: "October 7, 2017"
output: html_document
---
##Load Packages
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```


##H1B Data

Load the Data
```{r}
h1b <- read.csv(url("https://raw.githubusercontent.com/mikegankhuyag/607-Projects/master/Project%202/2007_2017_H1B_trend.csv"), stringsAsFactors = FALSE)
str(h1b)
View(h1b)
```

I'm particulary interested in which continent has the most h1b's. So lets take the data containing countries.

```{r}
Country <- data.frame(h1b[18:37,])
colnames(Country) <- c("Countries",2007,2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,"Total")
head(Country)
```

Lets add a column identifying the correct continent for each row.

```{r}
Country$Continent <- c("Asia", "Asia","Asia","Asia","North America","Asia","North America","Europe","Asia","Europe","South America","Asia","Asia","Europe","Europe","Asia","Europe","Asia","South America","Europe")

Country
```

I want the numbers to read as numbers and I want to remove the commas.

```{r}
cont <- select(Country,Countries,Continent,8:12)
cont$`2013` <- as.numeric(gsub(",","",cont$`2013`))
cont$`2014` <- as.numeric(gsub(",","",cont$`2014`))
cont$`2015` <- as.numeric(gsub(",","",cont$`2015`))
cont$`2016` <- as.numeric(gsub(",","",cont$`2016`))
cont$`2017` <- as.numeric(gsub(",","",cont$`2017`))
cont
```

Lets group the data by the continents we created.

```{r}
cont$Total5yrs = rowSums(cont[3:7])

h1b_continents <- summarise(group_by(cont,Continent),totalvisas = sum(Total5yrs))
h1b_continents



```

Create a visual graph containing the data.

```{r}
ggplot(h1b_continents, aes(x="", y=totalvisas, fill=Continent)) +
  geom_bar(width = 1, stat = "identity")

ggplot(h1b_continents, aes(x="", y=totalvisas, fill=Continent)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0)
```

##Small Business Data



Loading the data set and obserrving the structure.

```{r}
Businesses <- read.csv(url("https://raw.githubusercontent.com/mikegankhuyag/607-Projects/master/Project%202/1988_2014_payroll_firmsize.csv"))
str(Businesses)
```

Get rid of the Empty rows.

```{r}
B2 <- Businesses[-c(1:4,32,60,88,116,122:132),-c(14:16) ]
B2
```

Seperate the 4 tables consolidated together.

```{r}
Firms <- data.frame(B2[1:27,2:13])
Establishment <- data.frame(B2[28:54,2:13])
Employment <- data.frame(B2[55:81,2:13])
Payroll <- data.frame(B2[82:108,2:13])
```

Column names for four tables
```{R}
colnames(Firms) <- c("Year", "Non-employers",	"Employer Totals",	"0*",	"Less than 4"	,"5 to 9",	"10 to 19"	,"Less than 20",	"20 to 99",	"100 to 499",	"Less than 500",	"over 500")
colnames(Establishment) <- c("Year", "Non-employers",	"Employer Totals",	"0*",	"Less than 4"	,"5 to 9",	"10 to 19"	,"Less than 20",	"20 to 99",	"100 to 499",	"Less than 500",	"over 500")
colnames(Employment) <- c("Year", "Non-employers",	"Employer Totals",	"0*",	"Less than 4"	,"5 to 9",	"10 to 19"	,"Less than 20",	"20 to 99",	"100 to 499",	"Less than 500",	"over 500")
colnames(Payroll) <- c("Year", "Non-employers",	"Employer Totals",	"0*",	"Less than 4"	,"5 to 9",	"10 to 19"	,"Less than 20",	"20 to 99",	"100 to 499",	"Less than 500",	"over 500")
```

Change the order of the datasets.

```{r}
Firms <- arrange(Firms, -desc(Year))
Establishment <- arrange(Establishment, -desc(Year))
Employment <- arrange(Employment, -desc(Year))
Payroll <- arrange(Payroll, -desc(Year))
head(Firms)
head(Establishment)
head(Employment)
head(Payroll)
```



```{r}
library(lubridate) # for working with dates
library(ggplot2)  # for creating graphs
library(scales)   # to access breaks/formatting functions
library(gridExtra) # for arranging plots
```


```{r}
qplot(x=Firms$Year, y=Firms$`Less than 4`,
      data=Firms, na.rm=TRUE,
      main="Number of 0-4 Size Firms",
      xlab="Year", ylab="Number of Firms")
```

```{r}
ggplot(Firms, aes(Firms$Year, Firms$`Less than 4`)) +
  geom_point(na.rm = TRUE, color = "red") +
  ggtitle("Number of Small Businesses with Less than 4 Employers")+
  xlab("Year") + ylab("Number of Businesses")

```

```{r}
Less_than_4 <- ggplot()+
  geom_point(data = Firms, aes(Firms$Year, Firms$`Less than 4`), color = "red") +
  geom_point(data = Employment, aes(Employment$Year, Employment$`Less than 4`), color = "blue") +
  geom_point(data = Establishment, aes(Establishment$Year, Establishment$`Less than 4`), color = "green")+
  ggtitle("Number of Small Businesses with Less than 4 Employers")+
  xlab("Year") + ylab("Number of Businesses")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


Less_than_4 + scale_y_discrete(breaks = c(0,1,5000,5500,6000,10000))

```

```{r}
five_9 <- ggplot()+
  geom_point(data = Firms, aes(Firms$Year, Firms$`5 to 9`), color = "red") + 
  geom_point(data = Employment, aes(Employment$Year, Employment$`5 to 9`), color = "blue") + 
  geom_point(data = Establishment, aes(Establishment$Year, Establishment$`5 to 9`), color = "green")+ 
  geom_point(data = Payroll, aes(Payroll$Year, Payroll$`5 to 9`), color = "black")+
  ggtitle("Number of Small Businesses with 5 to 9 Employers")+
  xlab("Year") + ylab("Number of Businesses")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))




five_9 
```

```{r}
Small_business <- bind_cols(Firms,Employment, Establishment)

str(Small_business)
```
```{r}

L4 <- Small_business[,c(1,5,17,29)]
colnames(L4) <- c("Year","Firms","Employer","Establishments")
L4
```

```{r}
 ggplot()+
  geom_point(data = L4, aes(L4$Year, L4$Employer), color = "blue", col = "Employer") +
  geom_point(data = L4, aes(L4$Year, L4$Establishment), color = "green")+
  geom_point(data = L4, aes(L4$Year, L4$Firms), color = "red") +
  ggtitle("Number of Small Businesses with Less than 4 Employers")+
  xlab("Year") + ylab("Number of Businesses")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


```

```{r}
ggplot(L4, aes(L4$Year, y= value , color = variable))+
  geom_point(aes(y = L4$Employer, col = "Employers"))+
  geom_point(aes(y = L4$Establishments, col = "Establishments"))+
  geom_point(aes(y = L4$Firms, col = "Firms"))+
   ggtitle("Number of Small Businesses with Less than 4 Employers")+
  xlab("Year") + ylab("Number of Businesses")+
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


```

##NYC Collisions

Load the data

```{r}
collision <- read.csv(url("https://raw.githubusercontent.com/mikegankhuyag/607-Projects/master/Project%202/NYC_Collision.csv"), stringsAsFactors = FALSE)
head(collision)
str(collision)

```

I currently live in Manhattan, so I am most interested in typeo of collisions in Manhattan.


```{r}
collision2 <- data.frame(collision[6:27,])

collision2

```


```{r}
manhattan_collision <- t(data.frame(collision2[,2:14]))
colnames(manhattan_collision) <- manhattan_collision[1,]
manhattan_collision2 <- data.frame(manhattan_collision[2:13,])
manhattan_collision2
```

Change the data to numeric values

```{r}
manhattan_collision2$X1st.Precinct <-as.numeric(as.character( manhattan_collision2$X1st.Precinct))
manhattan_collision2$X5th.Precinct <-as.numeric(as.character( manhattan_collision2$X5th.Precinct))
manhattan_collision2$X6th.Precinct <-as.numeric(as.character( manhattan_collision2$X6th.Precinct))
manhattan_collision2$X7th.Precinct <-as.numeric(as.character( manhattan_collision2$X7th.Precinct))
manhattan_collision2$X9th.Precinct <-as.numeric(as.character( manhattan_collision2$X9th.Precinct))
manhattan_collision2$X10th.Precinct <-as.numeric(as.character( manhattan_collision2$X10th.Precinct))
manhattan_collision2$X13th.Precinct <-as.numeric(as.character( manhattan_collision2$X13th.Precinct))
manhattan_collision2$Midtown.South.Precinct <-as.numeric(as.character( manhattan_collision2$Midtown.South.Precinct))
manhattan_collision2$X17th.Precinct <-as.numeric(as.character( manhattan_collision2$X17th.Precinct))
manhattan_collision2$Midtown.North.Precinct <-as.numeric(as.character( manhattan_collision2$Midtown.North.Precinct))
manhattan_collision2$X19th.Precinct <-as.numeric(as.character( manhattan_collision2$X19th.Precinct))
  manhattan_collision2$X20th.Precinct <-as.numeric(as.character( manhattan_collision2$X20th.Precinct))
manhattan_collision2$Central.Park.Precinct <-as.numeric(as.character( manhattan_collision2$Central.Park.Precinct))
manhattan_collision2$X23rd.Precinct <-as.numeric(as.character( manhattan_collision2$X23rd.Precinct))
manhattan_collision2$X25th.Precinct <-as.numeric(as.character( manhattan_collision2$X25th.Precinct))
manhattan_collision2$X24th.Precinct <-as.numeric(as.character( manhattan_collision2$X24th.Precinct))
manhattan_collision2$X26th.Precinct <-as.numeric(as.character( manhattan_collision2$X26th.Precinct))
manhattan_collision2$X28th.Precinct <-as.numeric(as.character( manhattan_collision2$X28th.Precinct))
manhattan_collision2$X30th.Precinct <-as.numeric(as.character( manhattan_collision2$X30th.Precinct))
manhattan_collision2$X32nd.Precinct <-as.numeric(as.character( manhattan_collision2$X32nd.Precinct))
manhattan_collision2$X34th.Precinct <-as.numeric(as.character( manhattan_collision2$X34th.Precinct))
manhattan_collision2$X33rd.Precinct <-as.numeric(as.character( manhattan_collision2$X33rd.Precinct))
manhattan_collision2
```

Find the sum and the percentages

```{r}
manhattan_collision2$Total <- rowSums(manhattan_collision2[2:22])
Total_incidents <- sum(manhattan_collision2$Total)
manhattan_collision2$Percent <- (manhattan_collision2$Total/Total_incidents)

manhattan_collision2$Percent <- round(manhattan_collision2$Percent*100,digits = 2)
manhattan_collision2

```

Add a column for the type of collisions.

```{r}
manhattan_collision2$Type <- c("Number_of_Motor_Vehicle_Collisions","Vehicles_or_Motorists_Involved","Injury_or_Fatal_Collisions","MotoristsInjured",	"MotoristsKilled","PassengInjured","PassengKilled","CyclistsInjured","CyclistsKilled","PedestrInjured","PedestrKilled","Bicycle")

manhattan_collision3 <- manhattan_collision2 %>% select(Type,everything())
manhattan_collision3
```

Observe the data

```{r}

ggplot(manhattan_collision2, aes(x=Type, y=Total, fill=Type)) +
geom_bar(width = 1, stat = "identity")+
ggtitle("Manhattan Collisions in August")+
xlab("Type") + ylab("Amount")+
theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))

ggplot(manhattan_collision2, aes(x="", y=manhattan_collision3$Percent, fill=Type )) +
geom_bar(width = 1, stat = "identity")+
ggtitle("Percentage of Manhattan Collisions in August")+
xlab("") + ylab("Amount")+
theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))


```