---
title: "Project 1"
author: "Munkhnaran Gankhuyag"
date: "September 18, 2017"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
Chess Data.
```{r}
project1_data <- read.table("https://raw.githubusercontent.com/mikegankhuyag/607-Projects/master/tournamentinfo.txt",header = FALSE, sep = "\t", stringsAsFactors = FALSE)
head(project1_data)
library(stringr)
project1_data[1:10,]
                          
```

#The data for each player is in 2 lines, so I combined the first and second row of the player by finding the sequence for the lines.
```{r}
Chess_Data <- paste (project1_data[seq(5,196,3),],project1_data[seq(6,196,3),1],sep = "")
seq
head(Chess_Data)
```

#Where needed, I created seprators using the same '|' mark that the data uses. Also, I decided to input an "end" to mark where data becomes is irrelivant.

```{r}
Chess_data1 <- str_replace_all(Chess_Data, pattern = "\\:\\S?","|")
str_sub(Chess_data1,130, 134) <- "|end"
str_sub(Chess_data1,119, 120) <- " |"
head(Chess_data1)
```
#Using the | as a seperator, I created a table.

```{r}
Chess_data2 <-read.table(text = Chess_data1, sep = "|")
head(Chess_data2)
```

#Create a data frame out of the new table and name the columns.
```{r}
chess_data3 <- data.frame(Chess_data2[,1:14])
colnames(chess_data3) <-c("ID","Player","Score","Game1","Game2","Game3","Game4","Game5","Game6","Game7","State","USCF_ID","Pre_Rating", "Post_Rating")

chess_data3[1:10,]

```
#Clean out the data frame.
```{r}
chess_data3$USCF_ID <- unlist(str_extract_all(chess_data3$USCF_ID,"\\(?\\d{8}?"))
chess_data3$Pre_Rating <- substr(chess_data3$Pre_Rating, 1, 5)
chess_data3$Post_Rating <- substr(chess_data3$Post_Rating, 1, 4)

chess_data3[1:10,]

```
#Create a new data frame with just the opponent number and not the outcome of the match

```{r}
Chess_data_4 <- chess_data3

Chess_data_4$Game1 <- unlist((str_extract(Chess_data_4$Game1, "\\d+")))
Chess_data_4$Game2 <- unlist((str_extract(Chess_data_4$Game2, "\\d+")))
Chess_data_4$Game3 <- unlist((str_extract(Chess_data_4$Game3, "\\d+")))
Chess_data_4$Game4 <- unlist((str_extract(Chess_data_4$Game4, "\\d+")))
Chess_data_4$Game5 <- unlist((str_extract(Chess_data_4$Game5, "\\d+")))
Chess_data_4$Game6 <- unlist((str_extract(Chess_data_4$Game6, "\\d+")))
Chess_data_4$Game7 <- unlist((str_extract(Chess_data_4$Game7, "\\d+")))
head(Chess_data_4)
```

#Create a matrix from the data frame of just the ID and the pre rank of the opponent.

```{r}

Chess_data_Ranks <- matrix(c(Chess_data_4$ID, Chess_data_4$Pre_Rating), ncol = 2,nrow = 64)
colnames(Chess_data_Ranks) <- c("ID", "Pre_Rank")
Chess_data_Ranks <- data.frame(Chess_data_Ranks)

head(Chess_data_Ranks)

```
#Match opponent ID with their Pre rank scores and create a new tablae with just opponents pre rank averages.


```{r}
Chess_data_Ranks2 <- Chess_data_4

Chess_data_Ranks2$Game1 <- Chess_data_Ranks$Pre_Rank[ match(Chess_data_Ranks2$Game1, Chess_data_Ranks$ID)]
Chess_data_Ranks2$Game2 <- Chess_data_Ranks$Pre_Rank[ match(Chess_data_Ranks2$Game2, Chess_data_Ranks$ID)]
Chess_data_Ranks2$Game3 <- Chess_data_Ranks$Pre_Rank[ match(Chess_data_Ranks2$Game3, Chess_data_Ranks$ID)]
Chess_data_Ranks2$Game4 <- Chess_data_Ranks$Pre_Rank[ match(Chess_data_Ranks2$Game4, Chess_data_Ranks$ID)]
Chess_data_Ranks2$Game5 <- Chess_data_Ranks$Pre_Rank[ match(Chess_data_Ranks2$Game5, Chess_data_Ranks$ID)]
Chess_data_Ranks2$Game6 <- Chess_data_Ranks$Pre_Rank[ match(Chess_data_Ranks2$Game6, Chess_data_Ranks$ID)]
Chess_data_Ranks2$Game7 <- Chess_data_Ranks$Pre_Rank[ match(Chess_data_Ranks2$Game7, Chess_data_Ranks$ID)]

head(Chess_data_Ranks2)

```
#Turn all values in the game to a numeric value to cacluclate new rank

```{r}
Chess_data_Ranks2$Game1 <- as.numeric(as.character(Chess_data_Ranks2$Game1))
Chess_data_Ranks2$Game2 <- as.numeric(as.character(Chess_data_Ranks2$Game2))
Chess_data_Ranks2$Game3 <- as.numeric(as.character(Chess_data_Ranks2$Game3))
Chess_data_Ranks2$Game4 <- as.numeric(as.character(Chess_data_Ranks2$Game4))
Chess_data_Ranks2$Game5 <- as.numeric(as.character(Chess_data_Ranks2$Game5))
Chess_data_Ranks2$Game6 <- as.numeric(as.character(Chess_data_Ranks2$Game6))
Chess_data_Ranks2$Game7 <- as.numeric(as.character(Chess_data_Ranks2$Game7))
```

#Create a new column calculating the 

```{r}
New_Rating <- round(rowMeans(Chess_data_Ranks2[,4:10],na.rm = TRUE),digits = 0)

Chess_data_Ranks2$New_Rating <- New_Rating

Chess_data_Ranks2[,c(1,2,13,15)]



```

```{r}

Chess_data_Ranks2$Pre_Rating <- as.numeric(as.character(Chess_data_Ranks2$Pre_Rating))
Ranking_differences <- Chess_data_Ranks2$New_Rating -  Chess_data_Ranks2$Pre_Rating

Chess_data_Ranks2$Ranking_differences <- Ranking_differences

head(Chess_data_Ranks2)

summary(Chess_data_Ranks2$Ranking_differences)

Chess_data_Ranks2

```






